|

### 2024-12-19 Drafts/Paper_1/05-discussion.Rmd
Some discussion text

There are some limitations to this kind of life table reconstruction. Given we use historic data sets, the nature of the population growth are somewhat constrained, for example, when the stock is at the highest abundance in the observed time series, $\lambda$, by definition, has to be $\le$ 1, conversely at the lowest abundance observed $\lambda$ must be $\ge$ 1. There are more subtle constraints on the lambda estimates, for example, $\lambda$ must be $\le$ 2 when a stock is at 50% of the maximum abundance, and the minimum value of $\lambda$ similarly constrained by the minimun observed abundance observed in the time series. These constraints contributed to an increase in the variability of $lambda$ observed in the analysis.  In addition we assumed that the fishery removals from the stock assessments were correct and modified the productivity parameters to match the time series from the stock assessment; this resulted in $\lambda$ values that were 5% lower than expected using the initial productivity parameters from the stock assessment. If this decline in $\lambda$ was instead attributed to fishing mortality, then the underlying productivity of these systems would have resulted in less conservative projections (DOUBLING TIME... KEEP PULLING ON THIS THEAD
)
NOTE THAT THE BECAUSE WE ASSUMED FISHERY REMOVALS WERE CORRECT COMBINED WITH THE PRODUCITIVTY PARAMETERS HAVING RESULTED IN LAMBDA VALUES THAT ARE 5% LOWER, IF THAT DECLINE IN LAMBDA WAS DUE TO FISHING, OUR PROJECTIONS, GENERATION TIME, AND R0 VALUES WOULD ALL BE SOMEWHAT CONSERVATIVE AS TO GET THAT LOWER LAMBDA VALUE WE ATTRIBUTED THIS TO THE PRODUCTIVITY OF THE SYSTEM BEING LOWER, WHEREAS ATTRIBUTING THAT DIFFERENCE SOLELY TO FISHING WOULD HAVE HAD NO IMPACT ON THE NATURAL MORTALITY OR FECUNDITY

SOMETHING ABOUT EXPLOITATION HISTORY IF THERE ISN'T ALREADY.
MISSING CAVIATS (e.g. exploitation history, assumptions related to assuming model fit and fishing is right, )

>>>>

There are some limitations to this kind of life table reconstruction. Given we use historic data sets, the nature of the population growth are somewhat constrained, for example, when the stock is at the highest abundance in the observed time series, $\lambda$, by definition, has to be $\le$ 1, conversely at the lowest abundance observed $\lambda$ must be $\ge$ 1. There are also more subtle constraints on the lambda estimates, for example, $\lambda$ must be $\le$ 2 when a stock is at 50% of the maximum abundance, and the minimum value of $\lambda$ is similarly constrained by the minimum observed abundance observed in the time series. These constraints contributed to an increase in the variability of $lambda$ as abundance declines as observed in the analysis. In addition, we assumed that the fishery removals from the stock assessments were correct and modified the productivity parameters to match the time series from the stock assessment; this resulted in $\lambda_{F=0}$ values that were 5% lower than would have been calculated using the initial productivity parameters found in the stock assessments. If this decline in $\lambda$ was instead attributed to fishing mortality, then the underlying productivity of these systems would have resulted less conservative projections in the absence of fishing (e.g., the doubling time would have increased)... KEEP PULLING ON THIS THEAD

SOMETHING ABOUT EXPLOITATION HISTORY IF THERE ISN'T ALREADY.

WRAP UP PARA...
While most stocks exhibited a latent population growth capacity in the absence of fishing, regional and stock specific factors revealed that several stocks lacked this capacity, even in the absence of fishing. Knowledge of the underlying demographic parameters for a stock can be used to help sustainably manage the stock and can inform the development of rebuilding plans for stocks that have collapsed.  Finally, we note that life history characteristics alone were a poor predictor of annual population growth or future recovery potential, with greater variability observed among stocks rather than among species with distinct regions associated with generally higher or generally lower vital rates.
